---
title: "R Notebook"
output: html_notebook
---
```{r}
library(StereoMorph)
library(tidyverse)
library(geomorph)
library(listviewer)
library(stringr)
```

```{r}
(localities<-read_csv("Marks_Mox - Sheet1.csv"))
```
```{r}
localities %>%
  select(CatalogNumber,species,drainage,river)%>%
  rename(cat_num=CatalogNumber)->loc

```


```{r}
#files<-paste0("Mox_images/shapes/TU198404_",1:6,"_L.txt")
a<-readShapes(file = "Mox_images/shapes/",fields=c("landmarks.scaled","curves.scaled"))
```

Put landmark data into a tibble. 
```{r}
data<-tibble(num=attr(a$landmarks.scaled,"dimnames")[[3]],
       fixedlm=array_branch(a$landmarks.scaled,margin = 3),
       c_body_ant=map(a$curves.scaled,"body_ant"),
       c_body_post=map(a$curves.scaled,"body_post"),
       c_opercle=map(a$curves.scaled,"opercle"))

```

```{r}
data %>%
  mutate(cat_num=str_extract(num,"[0-9]+"))->data
```

```{r}
data$cat_num<-as.integer(data$cat_num)
(data<-left_join(data,loc))
```

convert curves into landmarks, evenly spaced along curves. Bind fixed and semi-landmarks together and remove duplicates. 
```{r}
data %>%
  mutate(body_ant=map(c_body_ant,~pointsAtEvenSpacing(.x,n=10)))%>%
  mutate(body_post=map(c_body_post,~pointsAtEvenSpacing(.x,n=10)))%>%
  mutate(opercle=map(c_opercle,~pointsAtEvenSpacing(.x,n=5))) %>%
  mutate(land_marks=pmap(list(fixedlm,body_ant,body_post,opercle),rbind))%>%
  mutate(land_marks=map(land_marks,~unique(.x)))->data
```

Convert list (and bind several arrays together) using `sapply()`
```{r}
new_a<-sapply(data$land_marks, I, simplify="array")
```


Generate plot to aid in defining sliding, semi-landmarks using AUTO mode of `define.sliders()`. There are 20 fixed landmarks, curves are found between landmarks 1 and 2 (anterior, dorsal body), 3 and 4 (posterior, dorsal body), and 10 and 11 (opercle) 

```{r}
dd<-as.data.frame(new_a[,,1])
dd$label<-1:length(dd$V1)

ggplot(dd,aes(V1,V2))+
  geom_point(alpha=0.7)+
  geom_text(label=dd$label,check_overlap = F,nudge_x = 1,size=3)
```

Generate semi-landmarks matrix for gpagen using `define.sliders()`. 
```{r}
curves<-rbind(define.sliders(c(1,29:36,2)),
              define.sliders(c(3,21:28,4)),
              define.sliders(c(10,37:39,11)))
```

```{r}
gpa<-gpagen(new_a,curves = curves)
gpa
```
```{r}
plotAllSpecimens(gpa$coords,mean=F)
```
plot PCA
```{r}
PCA <- plotTangentSpace(gpa$coords,warpgrids = F)
```
```{r}
pca<-as_tibble(PCA$pc.scores[,c(1,2)])
pca$id<-data$drainage
pca$id2<-data$species


pca %>%
  split(.$id)%>%
  map(~.x[chull(x=.x$PC1,y=.x$PC2),]) %>%
  bind_rows(.) -> hull
```



```{r}
ggplot(pca, aes(PC1, PC2, color = id2)) +
  geom_polygon(data = hull,aes(PC1,PC2,group=id,fill=id),alpha=0.3) +
  geom_point()+
  facet_wrap(~id2)
```

```{r}
M<-mshape(gpa$coords)
PC<-PCA$pc.scores[,1:2]
preds<-shape.predictor(gpa$coords,x=PC,Intercept = FALSE,
                       pred1=c(-0.05,0.04)) 
                       
GP<-gridPar(pt.size=0.5,tar.pt.size=0.5,n.col.cell = 50)      
plotRefToTarget(M,preds$pred1,mag = 2,method = "vector",gridPars = GP)
```



