---
title: "R Notebook"
output: html_notebook
---

Load packages needed for analysis
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
library(stringr)
```

load Bob Jenkins data. Includes counts and measurements that he collected between 1970-ish and 1990-ish. Included most, if not all, specimens in musuems at the time. 
```{r,cache=T}
mer_table<-read_csv(file = "AUSCON.csv")
mer_table
```
**Lateral Line Scales**

Summraise lateral line scales by river system (DRAINAGE) using a figure. Size of points is the number of individuals (n). The light grey point represents the mean value.  
  
Bob Jenkins thought, and my own analysis confirms (Clements et al., 2014), that it is important to consider variation within major drainages. Add some of that complexity to the plots and visualize. Size of points is the number of individuals (n). The light grey point represents the mean value.

```{r,fig.height=5}
mer_table %>%
  filter(LL_SC>0) %>%
  unite(drain_sys,DRAINAGE,SYSTEM,sep="_") %>%
  ggplot(aes(x=drain_sys,y=LL_SC))+
  geom_count() +
  theme(axis.text.x=element_text(angle=0,hjust=1,vjust = 1))+
  stat_summary(fun.y="mean",geom="point",colour = "grey", size = 2,alpha=.7)+
  scale_y_continuous(breaks=seq(41,50,1))+
  coord_flip()+
  ggtitle("Lateral Line Scales")
```
Fish from the Middle Rio Bravo (Alimito Creek, Rio Bravo Mainstem near the Conchos and Conchos proper) have higher lateral line scale counts (>45). Not too suprising given the genetic divergence (Clements et al. 2014) of this population  
  
The Rio Mezquital and San Pedro?_Lerma? (i.e.,*Scartomyzon milleri*) inhabit a separate, Pacific slope drainage and have similar counts as the Rio Conchos populations. 
  
ower Rio Grande de Santiago (Balanos, Juchipila, San Nichols, Huyamamato, Verde) Pacific slope drainages have a high lateral line counts, but lower mean counts compared to the Conchos and Rio Mezquital populations.  
  
Fish from above Lake Chapala (Lerma = type Locality for *S. austrinus*) have fewer lateral line scales (42-45) similar to populations in the Pacific slope Ameca and Armeria drainges. The latter fish represent the nominal *S. mascotae*.  
  
*S. albidus* are known from the tribs of Lower Rio Bravo (Rio San Juan, Rio Salado) and 1 Atlantic slope (Rio San Fernando) drainage of Mexico. All of these populations have counts (42-47) similar to *S. austrinus* and *S. mascotae* of the Pacific Slope.  
  
The undescribed Rio Soto la Marina populatio, formerly recognized as *S. albidus*, is genetically distinct from all other *Scartomyzon* in Mexico and Texas (Clements et al. 2014). This populations has fewer average lateral line scales (42-46) than *S. albiudus*. The counts of this population tend are simialar *S. congestus* (41-47) from the Atalantic drainges (Brazos, Colorado, Guadelupe, San Antonio, Nueces). The Pecos R., recognized as *S. congestus* (Clements et al., 2014), have similar counts (44-45).
  
The fish from Devils R and vicinity were genetically similar to *S. congestus* (Clements et al.  2014).These fish tend to have higher counts (44-47) than other Atlantic slope and Pecos R *S. congestus* populations.They are more similar to fish from the lower Rio Bravo (Rio San Juan and Rio Salado).  
  
  
  
**Circumbody Scales**  
Summraise circumbody scales by river system (DRAINAGE and system) with a figure. Size of points is the number of individuals (n). The light grey point represents the mean value. Include system level information for certain drainages. 

```{r,fig.width=8}
mer_table %>%
  filter(CB_SC>0) %>%
  unite(drain_sys,DRAINAGE,SYSTEM,sep="_")%>%
  
  ggplot(aes(x=drain_sys,y=CB_SC))+
  geom_count() +
  theme(axis.text.x=element_text(angle=0,hjust=1,vjust = 1))+
  stat_summary(fun.y="mean",geom="point",colour = "grey", size = 2,alpha=.7)+
  scale_y_continuous(breaks=seq(30,44,1))+
  coord_flip()+
  ggtitle("Total Circumbody Scales")
```
Similar to the patterns seen in lateral line scale counts. Conchos River (38-43) and Lower Rio Gande de Santiago (36-43) have high sclae counts.  
  
Unlike the pattern in lateral line scales, the Rio Mezquital (*S. milleri*) have fewer circumbody scales (35-38) relative to Conchos and lower Rio Grande de Santiago.  
  
The Lerma (*S. austrinus, type locality) have the lowest counts (30-34).   
  
*S. moscotae* (Aremeria and Ameca) have more circumbody scales (32-39) than the nearby Rio Lerma. 
  
*S.albiuds* (Lower Rio Bravo tribs and Atlantic slope draingages of Mexico) have 31-37 scales. The Rio Salado tends to have lower counts (31-33). The undescribe Rio Soto la Marina have simialr counts. Circumbody scales are similar for *S. congestus* (31-37) (Pecos R, Devils R, and Atlantic Slope Drainages of Texas).   
  
  
  
**Upper Circumbody Scales** 
```{r,fig.width=7}
mer_table %>%
  filter(CB_U>0)%>%
  unite(drain_sys,DRAINAGE,SYSTEM,sep="_")%>%
  ggplot(aes(x=drain_sys,y=CB_U))+
  geom_count() +
  theme(axis.text.x=element_text(angle=0,hjust=1,vjust = 1))+
  stat_summary(fun.y="mean",geom="point",colour = "grey", size = 3,alpha=.7)+
  scale_y_continuous(breaks=seq(10,50,1))+
  coord_flip()+
  ggtitle("Upper Circumbody Scales")


```
